---
title: "Homework 8"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---
UT EID: kt27955
Github Link: https://github.com/ktaspa/Homework8SDS315

## Problem 1: Regression Warm-Up

### A)

We fit a linear regression model with creatclear as the response and age
as the predictor. Then with the model we get the value 113.723 for the
creatinine clearance rate we expect for a 55-year-old.

```{r, echo = FALSE}
creatinine <- read.csv("creatinine.csv")
model <- lm(creatclear ~ age, data = creatinine)
predict(model, newdata = data.frame(age = 55))
```
### B)

The rate of change of creatinine clearance is negative so it decreases
with age. We get the slope of the linear regression.

```{r, echo = FALSE}
summary(model)$coefficients["age", "Estimate"]
```
### C)

Compared the actual creatinine clearance rates of two patients to the
expected values from the model and got the 40-year-old with a rate of
135 having a higher creatinine clearance rate. Equation: observed -
predicted is difference between the two groups.

```{r, echo = FALSE}
pred_40 <- predict(model, newdata = data.frame(age = 40))
pred_60 <- predict(model, newdata = data.frame(age = 60))
obs_40 <- 135
obs_60 <- 112
diff_40 <- obs_40 - pred_40
diff_60 <- obs_60 - pred_60
c(diff_40, diff_60)
```

## Problem 2: Modeling Disease Growth

### 1.

The growth rate is 0.183 and the doubling time is 3.8.\
The growth rate confidence interval with 95% confidence is (0.162,
0.204)\
The doubling time confidence interval with 95% confidence is (3.4, 4.3)\
Both intervals are positive and go not contain 0, meaning that both
values are statistically significant.

```{r, echo = FALSE}
covid <- read.csv("covid.csv")
italy <-subset(covid, country == "Italy")
spain <-subset(covid, country == "Spain")

italy_model <- lm(log(deaths) ~ days_since_first_death, data = italy)

coef(italy_model)["days_since_first_death"]
log(2) / (coef(italy_model)["days_since_first_death"])

confint(italy_model)["days_since_first_death", ]
log(2) / rev(confint(italy_model)["days_since_first_death", ])
```
### 2.

The growth rate is 0.276 and the doubling time is 2.5.\
The growth rate confidence interval with 95% confidence is (0.237,
0.316)\
The doubling time confidence interval with 95% confidence is (2.2,2.9)\
Both intervals are positive and go not contain 0, meaning that both
values are statistically significant.

```{r, echo = FALSE}
spain_model <- lm(log(deaths) ~ days_since_first_death, data = spain)

coef(spain_model)["days_since_first_death"]
log(2) / (coef(spain_model)["days_since_first_death"])

confint(spain_model)["days_since_first_death", ]
log(2) / rev(confint(spain_model)["days_since_first_death", ])
```
### 3.

```{r, echo = FALSE, warning = FALSE}
library(ggplot2)

ggplot(covid, aes(x = days_since_first_death, y = deaths, color = country)) +
  geom_line(size = 1.1) +
  labs(
    title = "Daily COVID-19 Deaths in Italy vs Spain",
    x = "Days Since First Death",
    y = "Number of Deaths",
    color = "Country"
  )
```

## Problem 3: Price Elasticity of Demand

The price elasticity of milk is -1.619. To estimate the elasticity I used a model that compares the percent change in sales to the percent change in price and the slope of that line tells us the price elasticity.
```{r, echo=FALSE}
milk <- read.csv("milk.csv")
milk$log_price <- log(milk$price)
milk$log_sales <- log(milk$sales)

elasticity_model <- lm(log_sales ~ log_price, data = milk)
coef(elasticity_model)["log_price"]

```
Confidence interval done with confint at the 95% confidence level gave an interval of (-1.779, -1.458) the interval is all negative and does not contain 0 which means that the price lelasticity of milk is statistically significant. 
```{r, echo = FALSE}
confint(elasticity_model)["log_price", ]
```
